---
title: "RFBO Trips"
---

```{r}
#| include: false
library(moveHMM)
library(targets)
library(tidyverse)

# Specifying the store makes "Run Document" work
hmm <- tar_read("rfbo_hmm", store = here::here("_targets"))
tracks <- tar_read("rfbo_decoded", store = here::here("_targets"))

palette = c(rest = "#E4A117",
            search = "#61B6E7",
            transit = "#1BA179",
            total = "black")
linetypes = c(rest = "solid",
              search = "solid",
              transit = "solid",
              total = "dashed")

colony_distance_km <- 20
duration_hr <- 2
max_gaps <- 4
```

## Trip summary

RFBO tracks were cleaned and filtered according to the following:

-   Tracks were split into trips, defined as excursions \>`r colony_distance_km` km from the colony lasting \>`r duration_hr` hours.

-   Relocations within trips were re-discretized using linear interpolation at 2 min intervals.

-   Only trips with gaps ≤`r max_gaps` relocation`r ifelse(max_gaps == 1, "", "s")` were retained.

That left `r n_distinct(tracks$burst)` trips from `r n_distinct(tracks$id)` individuals. This sample size is small, so we probably want to relax the constraints.

## Movement modeling

A 3-state hidden Markov model was fit to the movement data (step lengths and turning angles) to classify movement behavior as transit, search, and rest.

### Initial parameter selection

The initial parameters were estimated via mixture modeling. The step length distribution was assumed to come from a mixture of three gamma distributions.

| State     | $\mu$ | $\sigma$ |
|-----------|-------|----------|
| *transit* | 1.3   | 0.30     |
| *search*  | 0.38  | 0.18     |
| *rest*    | 0.049 | 0.022    |

: Initial step distribution parameters (in km).

Likewise, the turning angle distribution was assumed to come from a mixture of three von Mises distributions. von Mises parameters ($\mu$ and $\kappa$) were estimated from the relocations as classified by the step length mixture model.

| State     | $\mu$  | $\kappa$ |
|-----------|--------|----------|
| *transit* | -0.010 | 6.5      |
| *search*  | -0.011 | 1.3      |
| *rest*    | 0.0080 | 2.5      |

: Initial turning angle distribution parameters (in radians).

### HMM parameters

```{r}
#| label: hmm-param
#| fig-cap: Step length distribution from HMM

hmm_plot_data <- getPlotData(hmm, type = "dist", format = "long")
step_plot_data <- hmm_plot_data$step %>% 
  mutate(state = case_when(
    state == "state 1" ~ "rest",
    state == "state 2" ~ "search",
    state == "state 3" ~ "transit",
    TRUE ~ "total"
  ))

ggplot(step_plot_data, 
       aes(x = step, y = density, color = state, linetype = state)) +
  geom_line() +
  scale_color_manual(values = palette) +
  scale_linetype_manual(values = linetypes) +
  labs(x = "Step length (km)") +
  theme_classic() +
  theme(legend.position = c(1, 1),
        legend.justification = c(1, 1),
        legend.title = element_blank())
```

### Decoded trips

All the RFBO trips with behavioral state decoded using the Viterbi algorithm with the HMM. Gold point is KPNWR for reference.

```{r}
trips <- unique(tracks$burst)
kpnwr <- c(lon = -159.4, lat = 22.2)

for (trip in trips) {
  burst <- filter(tracks, burst == trip) %>% 
    mutate(state = factor(state, 
                          levels = 1:3, 
                          labels = c("rest", "search", "transit"))) %>% 
    drop_na(x, y)
  p <- ggplot(burst, aes(x, y, color = state, group = 1)) +
    geom_path() +
    geom_point() +
    annotate("point", 
             kpnwr["lon"], kpnwr["lat"], 
             color = "gold", size = 2, shape = 19) +
    scale_color_manual(values = palette) +
    labs(title = burst$burst[1]) +
    coord_map() +
    theme_classic() +
    theme(legend.position = "bottom",
          legend.title = element_blank())
  print(p)
}
```
