---
title: "RFBO Trips"
format: 
  html:
    theme: yeti
code-fold: true
---

```{r}
#| include: false
library(moveHMM)
library(targets)
library(tidyverse)

hmm <- tar_read("rfbo_hmm")
tracks <- tar_read("rfbo_decoded")
```

## Trip summary

RFBO tracks were cleaned and filtered according to the following:

-   Tracks were split into trips, defined as excursions \>50 km from the colony lasting \>2 hours.

-   Relocations within trips were re-discretized using linear interpolation at 2 min intervals.

-   Only trips with gaps ≤1 relocation were retained.

That left `r n_distinct(tracks$burst)` trips from `r n_distinct(tracks$id)` individuals. This sample size is small, so we probably want to relax the constraints.

## Movement modeling

A 3-state hidden Markov model was fit to the movement data (step lengths and turning angles) to classify movement behavior as transit, search, and rest.

### Initial parameter selection

The initial parameters were estimated via mixture modeling. The step length distribution was assumed to come from a mixture of three gamma distributions.

| State      | $\mu$ | $\sigma$ |
|------------|-------|----------|
| *transit*  | 1.3   | 0.37     |
| *search*   | 0.37  | 0.18     |
| *rest*     | 0.048 | 0.028    |

: Initial step distribution parameters (in km).

Likewise, the turning angle distribution was assumed to come from a mixture of three von Mises distributions. von Mises parameters ($\mu$ and $\kappa$) were estimated from the relocations as classified by the step length mixture model.

| State      | $\mu$  | $\kappa$ |
|------------|--------|----------|
| *transit*  | -0.011 | 7.2      |
| *search*   | -0.043 | 1.7      |
| *rest*     | -0.015 | 2.7      |

: Initial turning angle distribution parameters (in radians).

### HMM parameters

```{r}
#| label: hmm-param
#| fig-cap: Step length distribution from HMM

hmm_plot_data <- getPlotData(hmm, type = "dist", format = "long")
step_plot_data <- hmm_plot_data$step %>% 
  mutate(state = case_when(
    state == "state 1" ~ "rest",
    state == "state 2" ~ "search",
    state == "state 3" ~ "transit",
    TRUE ~ "total"
  ))

palette = c(rest = "#E4A117",
            search = "#61B6E7",
            transit = "#1BA179",
            total = "black")
linetypes = c(rest = "solid",
              search = "solid",
              transit = "solid",
              total = "dashed")

ggplot(step_plot_data, 
       aes(x = step, y = density, color = state, linetype = state)) +
  geom_line() +
  scale_color_manual(values = palette) +
  scale_linetype_manual(values = linetypes) +
  labs(x = "Step length (km)") +
  theme_classic() +
  theme(legend.position = c(1, 1),
        legend.justification = c(1, 1),
        legend.title = element_blank())
```

### Example decoded trip

Map of a trip with and without KPNWR for scale.

```{r}
#| label: decoded-trips

kpnwr <- c(lon = -159.4, lat = 22.2)

burst <- filter(tracks, burst == "1149_0046") %>% 
  mutate(state = factor(state, 
                        levels = 1:3, 
                        labels = c("rest", "search", "transit")))
ggplot(burst, aes(x, y, color = state, group = 1)) +
  geom_path() +
  geom_point() +
  annotate("point", 
           kpnwr["lon"], kpnwr["lat"], 
           color = "gold", size = 2, shape = 19) +
  scale_color_manual(values = palette) +
  coord_map() +
  theme_classic() +
  theme(legend.justification = c(0, 0),
        legend.position = c(0.01, 0.25),
        legend.title = element_blank())
ggplot(burst, aes(x, y, color = state, group = 1)) +
  geom_path() +
  geom_point() +
  scale_color_manual(values = palette) +
  coord_map() +
  theme_classic() +
  theme(legend.justification = c(0, 0),
        legend.position = c(0.45, 0.25),
        legend.title = element_blank())

```
